---
title: "Setting up for R tutorial"
author: "Kesson"
date: "16/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What you’ll need:
This lesson assumes you have current versions of the following software installed on your computer:

1.	the [R software](https://cran.r-project.org/) itself, and
2.	[RStudio Desktop](https://rstudio.com/products/rstudio/download/)
3.  the R package *tidyverse* (a package is a set of software tools that you use in R)

Both programmes and all R packages are free. RStudio runs the R software within it, so first download R package and install it on your computer. Then download and install RStudio. For detailed instructions for Macs or PCs, see <https://datacarpentry.org/r-socialsci/setup.html>

If you have never used R or RStudio before, I recommend going through [this tutorial](https://datacarpentry.org/r-socialsci/00-intro/index.html), it takes about 30 minutes. At the end of it you will download the package *tidyverse*

If you are familiar with RStudio install *tidyverse* by entering the command into your console:

`install.packages("tidyverse")`

## What we'll do

On Monday we will be going through the tutorial ["Manipulating, analyzing and exporting data with tidyverse"](https://datacarpentry.org/R-ecology-lesson/03-dplyr.html) 

This tutorial uses an example dataset we'll call surveys. To import this dataset run the following code:
```{r, echo=F, message=FALSE}
library(readr)
surveys <- read_csv("https://ndownloader.figshare.com/files/2292169")
```
This will create your first R object, a dataset called "surveys". All exercises will be run on this dataset.

## Useful stuff

There are tons of free resources for learning how to use R. 
For online resources I recommend: 
Cheatsheets:
The cheat sheets make it easy to learn about and use some of our favorite packages. They are published in their respective PDF versions here: <https://www.rstudio.com/resources/cheatsheets/>, some are also available in RStudio under Help-Cheatsheets.

For Monday's tutorial: 
*  [Data transformation with dplyr](https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf)
*  [RStudio guide](https://github.com/rstudio/cheatsheets/blob/master/rstudio-ide.pdf)

As a textbook:
[R for Data Science](https://r4ds.had.co.nz) by Hadley Wickham, the Buddah of the Tidyverse

For curriculum I base tutorials around the Data/Software Carpentries:
<https://datacarpentry.org/lessons/>
<http://swcarpentry.github.io/r-novice-inflammation/>